---
title: "Exam"
author: "Tanjina Islam, Miguel Morales Expósito and Carlos Perales Liñan, group 12"
date: "21 March 2018"
output:
  pdf_document:
    fig_caption: yes
fontsize: 11pt
highlight: tango
---
```{r,echo=FALSE, warning=FALSE,results='hide',message=FALSE}
options(digits = 3) #Showing only 3 decimals
library(multcomp)
library(lme4)
```


## Exercise 1




###Task 2

###Task 2

```{r}
galapagos=read.table("gala.txt", header=TRUE)

```


With this method, we start with all of the possible variables in our model. Then, we choose the one that gives the highest p-value. If this p-value is bigger than 0.05, we will discard the variable and repeat the process without it.

```{r}

galalm=lm(sqrt(Species)~Area+Elevation+Nearest+Scruz+Adjacent,data=galapagos) 
summary(galalm)[[4]]
```

The variable Nearest has the highest p-value with a p-value of [`r summary(galalm)[[4]][22]`]. Since this p-value is bigger than 0.05, we discard it for our model and continue to the next iteration.

```{r}
galalm=lm(sqrt(Species)~Area+Elevation+Scruz+Adjacent,data=galapagos) 
summary(galalm)[[4]]
```

The variable Scruz has the highest p-value with a p-value of [`r summary(galalm)[[4]][19]`]. Since this p-value is bigger than 0.05, we discard it for our model and continue to the next iteration.

```{r}
galalm=lm(sqrt(Species)~Area+Elevation+Adjacent,data=galapagos) 
summary(galalm)[[4]]
```

The variable Area has the highest p-value with a p-value of [`r summary(galalm)[[4]][14]`]. Since the p-value is bigger than 0.05, we discard it for our model and continue to the next iteration.

```{r}
galalm=lm(sqrt(Species)~Elevation+Adjacent,data=galapagos) 
summary(galalm)[[4]]
```

As we can see, all the p-values are smaller than 0.05, thus meaning that all the variables are significant for our model.


The resulting model of the step-down method is:
  
  sqrt(Species) = 3.5379 + 0.012942\*Elevation -0.00289\*Adjacent + error
