---
title: "Exam"
author: "Tanjina Islam, Miguel Morales Expósito and Carlos Perales Liñan, group 12"
date: "21 March 2018"
output:
  pdf_document:
    fig_caption: yes
fontsize: 11pt
highlight: tango
---
```{r,echo=FALSE, warning=FALSE,results='hide',message=FALSE}
options(digits = 3) #Showing only 3 decimals
library(multcomp)
library(lme4)
```


## Exercise 1



###Task 2

```{r}
galapagos=read.table("gala.txt", header=TRUE)

```


With this method, we start with all of the possible variables in our model. Then, we choose the one that gives the highest p-value. If this p-value is bigger than 0.05, we will discard the variable and repeat the process without it.

```{r}

galaglm=glm(sqrt(Species)~Area+Elevation+Nearest+Scruz+Adjacent,family=poisson,data=galapagos)
summary(galaglm)
```

The variable Nearest has the highest p-value with a p-value of 0.411. Since this p-value is bigger than 0.05, we discard it for our model and continue to the next iteration.

```{r}
galaglm=glm(sqrt(Species)~Area+Elevation+Scruz+Adjacent,family=poisson,data=galapagos) 
summary(galaglm)
```

The variable Scruz has the highest p-value with a p-value of 0.2466. Since this p-value is bigger than 0.05, we discard it for our model and continue to the next iteration.

```{r}
galaglm=glm(sqrt(Species)~Area+Elevation+Adjacent,family=poisson,data=galapagos) 
summary(galaglm)
```

As we can see, all the p-values are smaller than 0.05, thus meaning that all the variables are significant for our model.


The resulting model of the step-down method is:
  
sqrt(Species) = 1.314e+00 + -3.262e-04\*Area + 2.018e-03\*Elevation -3.987e-04\*Adjacent + error
