
---
title: "Assignment 4"
author: "Tanjina Islam, Miguel Morales Expósito and Carlos Perales Liñan, group 12"
date: "22 March 2018"
output:
  pdf_document:
    fig_caption: yes
fontsize: 11pt
highlight: tango
---
```{r,echo=FALSE, warning=FALSE,results='hide',message=FALSE}
options(digits = 3) #Showing only 3 decimals
library(multcomp)
library(lme4)
```


## Exercise 2

We load the data from the data source.
```{r}

```

### Task 1 



```{r}


```

#### Findings: 

### Task 2

```{r}


```

#### Findings: 

### Task 3

```{r}


```

#### Findings: 

### Task 4

```{r}


```

#### Findings: 


### Task 5

```{r}


```

#### Findings: 

### Task 6

```{r}


```

#### Findings: 

### Task 7

```{r}


```

#### Findings: 

### Task 8

```{r}


```

#### Findings: 

### Task 9

```{r}


```

#### Findings: 

### Task 10

```{r}


```

#### Findings: 


## Exercise 1

We load the data from the data source.
```{r}

```

### Task 1 



```{r}


```

#### Findings: 

### Task 2

```{r}


```

#### Findings: 

### Task 3

```{r}


```

#### Findings: 

### Task 4

```{r}


```

#### Findings: 


### Task 5

```{r}


```

#### Findings: 

### Task 6

```{r}


```

#### Findings: 

### Task 7

```{r}


```

#### Findings: 

### Task 8

```{r}


```

#### Findings: 


## Exercise 3

We load the data from the data source.
```{r}

```

### Task 1 



```{r}


```

#### Findings: 

### Task 2

```{r}


```

#### Findings: 

### Task 3

```{r}


```

#### Findings: 

### Task 4

```{r}


```

#### Findings: 


### Task 5

```{r}


```

#### Findings: 



