---
title: "Assignment 3"
author: "Tanjina Islam, Miguel Morales Expósito and Carlos Perales Liñan, group 12"
date: "13 March 2018"
output:
  pdf_document:
    fig_caption: yes
fontsize: 11pt
highlight: tango
---

## Exercise 1
Exercise 1
```{r}
# Add Code
```

### ( 1 ) 
Exercise 1: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 1: Task-2
```{r}
# Add Code
```

## Exercise 2
Exercise 2
```{r}
# Add Code
```

### ( 1 ) 
Exercise 2: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 2: Task-2
```{r}
# Add Code
```


### ( 3 ) 
Exercise 2: Task-3
```{r}
# Add Code
```

### ( 4 )
Exercise 2: Task-4
```{r}
# Add Code
```

## Exercise 3
Exercise 3
```{r}
# Add Code
```

### ( 1 ) 
Exercise 3: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 3: Task-2
```{r}
# Add Code
```

## Exercise 4
Exercise 4
```{r}
# Add Code
```

### ( 1 ) 
Exercise 4: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 4: Task-2
```{r}
# Add Code
```

### ( 3 )
Exercise 4: Task-3
```{r}
# Add Code
```

## Exercise 5
Exercise 5
```{r}
# Add Code
```

### ( 1 ) 
Exercise 5: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 5: Task-2
```{r}
# Add Code
```

## Exercise 6
Exercise 6
```{r}
# Add Code
```

### ( 1 ) 
Exercise 6: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 6: Task-2
```{r}
# Add Code
```


### ( 3 ) 
Exercise 6: Task-3
```{r}
# Add Code
```

### ( 4 )
Exercise 6: Task-4
```{r}
# Add Code
```

### ( 5 ) 
Exercise 6: Task-5
```{r}
# Add Code
```

### ( 6 )
Exercise 6: Task-6
```{r}
# Add Code
```

## Exercise 7
Exercise 7
```{r}
# Add Code
```

### ( 1 ) 
Exercise 7: Task-1
```{r}
# Add Code
```

### ( 2 )
Exercise 7: Task-2
```{r}
# Add Code
```


### ( 3 ) 
Exercise 7: Task-3
```{r}
# Add Code
```

## Our Explanations/Findings : 